Keyword and Collocation analysis on India-related text of Natural History

### 2023-06-21 paper_quarto.qmd
keyword wordlist update

Natural History by Pliny the Elder is considered the very first encyclopedia in modern history. It not only showcased the actual merchant exchange of the Roman Empire, but also manifested the spatial imagination of Pliny the Elder and the referenced scholarships of that time.

The text of the whole book has been digitized and annotated with people's name and places' name and coordinates by [TOPOSText](https://topostext.org/the-project) project since 2012. By scraping the text as well as the geographical related annotations, the distribution of place names in the encyclopedia can be investigated, which will contribute to a comprehensive understanding of the spatial imagination of Pliny the Elder and the scholarships in his reference network.

@fig-geoname shows the Top 20 distinct geographical names mentioned in Natural History.

```{r}
#| label: fig-geoname
#| fig-cap: Bar Chart for Top 20 Distinct Place Names Mentioned in Natural History
#| warning: false
#| echo: false

@@

When mapping the places with the annotated coordinates, as shown in @fig-map, India is the only one place that not in Mediterranean region while having a rather high counts in the mentioned times.

```{r}
#| label: fig-map
#| fig-cap: Mapping of geographical names mentioned in Natural History
#| warning: false
#| echo: false

@@

In this regards, this paper tries to further explore what themes are referred around the text about India region with application of keyword analysis and collocation analysis methods in corpus linguistics.

@@

As mentioned, the digitized text with geographical annotation is available through [TOPOSText](https://topostext.org/work/148 "Check Natural History on ToposText") project, I further scraped the full text and exported them in .txt as a **reference corpus**, and also exported those with annotation of coordinates in a range of Indian subcontinent as geographically related texts to India as a **target corpus**. The scraping code can be accessed in the link.

@@

The tokens are separated by white space and in the analysis of this paper, all the punctuation marks are removed.

[@R-tidyverse] and [@R-mclm] [@R-base] [@R-ggplot2] [@R-leaflet] are the main packages used in this paper. [@dodd2021] [@nappo2017]

@@

@@

# build frequency list for target corpus
flist_target <- fnames_indiatext %>%
  freqlist(
    re_token_splitter = r"--[(?xi)    \s+   ]--", # whitespace as token splitter
    re_token_transf_in = "[[:punct:]]", # Match punctuation marks
    token_transf_out = "" # Replace punctuation marks with an empty string
  )

# build frequency list for reference corpus
flist_ref <- fnames_wholetext %>%
  freqlist(re_token_splitter = r"--[(?xi)    \s+   ]--", # whitespace as token splitter
           re_token_transf_in = "[[:punct:]]", # Match punctuation marks
           token_transf_out = "")

@@

@tbl-kwtables below shows the top 30 keywords in Indian related texts comparing to other context, respectively sorted by PMI and G_signed.

::: {#tbl-kwtables .panel-tabset}
## Sorted by PMI

```{r}
#| warning: false
#| echo: false
#| label: tbl-kwtable_PMI
#| tbl-cap-location: top
#| tbl-cap: Top 30 Keywords on Indian Related Text in Natural History (sorted by PMI)

top_scores_kw %>% 
  as_tibble() %>%
  select(type, a, PMI, G_signed) %>% # select 4 columns
  arrange(desc(PMI)) %>%             # sort by PMI (descending) 
  head(30) %>%                       # select top 30 rows
  kbl(col.names = c("Type", "Frequency", "PMI", r"(Signed $G^2$)")) %>% 
  kable_minimal() %>% 
  scroll_box(height = "400px")
```

@@

```{r}
#| warning: false
#| echo: false
#| label: tbl-kwtable_Gsigned
#| tbl-cap-location: top
#| tbl-cap: Top 30 Keywords on Indian Related Text in Natural History (sorted by G_Signed)

top_scores_kw %>% # also valid for top_scores_colloc
  as_tibble() %>%
  select(type, a, PMI, G_signed) %>% # select 4 columns
  arrange(desc(G_signed)) %>%             # sort by PMI (descending) 
  head(30) %>%                       # select top 30 rows
  kbl(col.names = c("Type", "Frequency", "PMI", r"(Signed $G^2$)"), escape = FALSE) %>% 
  kable_minimal() %>% 
  scroll_box(height = "400px")
```
:::

```{r}
#| label: tbl-kwtables1
#| tbl-cap: Top 30 Keywords of Indian related texts in Natural History
#| tbl-subcap: 
#|  - "Sorted by PMI"
#|  - "Sorted by G_signed"
#| layout-ncol: 2
#| warning: false
#| echo: false
#| output: false

top_scores_kw %>% # also valid for top_scores_colloc
  as_tibble() %>%
  select(type, a, PMI, G_signed) %>% # select 4 columns
  arrange(desc(PMI)) %>%             # sort by PMI (descending) 
  head(30) %>%                       # select top 30 rows
  kbl(col.names = c("Type", "Frequency", "PMI", "r (Signed $G^{2}$)"))

top_scores_kw %>% # also valid for top_scores_colloc
  as_tibble() %>%
  select(type, a, PMI, G_signed) %>% # select 4 columns
  arrange(desc(G_signed)) %>%             # sort by PMI (descending) 
  head(30) %>%                       # select top 30 rows
  kbl(col.names = c("Type", "Frequency", "PMI", "r (Signed $G^{2}$)"))

>>>>

Pliny the Elder's "Natural History" stands as a seminal work in the history of encyclopedias, serving as a gateway to understanding the knowledge and worldview of ancient Rome. Beyond its role as a compendium of diverse subjects, the text provides a captivating glimpse into the actual merchant exchanges [@dodd2021] of the Roman Empire while shedding a light on the spatial imagination of Pliny and the scholarly references of his time [@nappo2017].

The text of the whole book has been digitized and annotated with people's name, places' name and coordinates by [TOPOSText](https://topostext.org/the-project) project since 2012. This invaluable resource allows for the creation of a dataset that includes both the textual contents and geographical annotations, which can be utilized to investigate the distribution of place names in the entire text and examine the frequencies and patterns of geographically-related content.

As a starting point, @fig-geoname reveals the Top 20 distinct geographical names mentioned within the text.

```{r}
#| label: fig-geoname
#| fig-cap: Bar Chart for Top 20 Distinct Place Names Mentioned in Natural History
#| fig-cap-location: top
#| warning: false
#| echo: false

@@

An intriguing observation from this analysis, as depicted in @fig-map, is the prominence of India---a region outside the Mediterranean---despite its high frequency of mentions.

```{r}
#| label: fig-map
#| fig-cap: Mapping of geographical names mentioned in Natural History
#| fig-cap-location: top
#| warning: false
#| echo: false

@@

Building upon this observation, this paper aims to delve deeper into the themes associated with the India region by employing keyword analysis and collocation analysis methods within the framework of corpus linguistics.

The analysis is conducted in R environment [@R-base], and the main corpus linguistics anlysing methods are applied through functions from [mclm](https://masterclm.github.io/mclm/) [@R-mclm] package. In addition, [tidyverse](https://www.tidyverse.org/) [@R-tidyverse] and [leaflet](https://rstudio.github.io/leaflet/) [@R-leaflet] are used for data manipulation and visualisation.

@@

As mentioned earlier, the digitized text of "Natural History" with geographical annotations is readily accessible through the [TOPOSText](https://topostext.org/work/148 "Check Natural History on ToposText") project. In order to facilitate the analysis, I conducted additional scraping of the full text, exporting it as a **reference corpus** in .txt format. Furthermore, I extracted the texts specifically associated with the Indian subcontinent, utilizing the annotated coordinates, creating a **target corpus** of geographically related texts focused on India. The scraping code can be found in the link: .

@@

The punctuation marks are removed in the tokenization process.

@@

The frequency list of the **target corpus** (Indian subcontinent related texts) and **reference corpus** (whole text) are generated respectively. And the association scores are calculated and filtered for those returning both PMI and $G^2$ gaining 2 or higher, in order to get the statisticly most significant keyword in India-related texts comparing to the reference corpus.

When calculating the word frequency in the **target corpus**, the word "india" itself is removed from the token in order to get more descriptive words in the corpus, which will be more efficient in implying the possible thematic keyword in the **target corpus**.

@@

# build frequency list for target corpus
flist_target <- fnames_indiatext %>%
  freqlist(
    re_token_transf_in = "[[:punct:]]", # Match punctuation marks
    token_transf_out = "", # Replace punctuation marks with an empty string
    re_drop_token = "india"
  )

# build frequency list for reference corpus
flist_ref <- fnames_wholetext %>%
  freqlist(
    re_token_transf_in = "[[:punct:]]", # Match punctuation marks
    token_transf_out = ""
  )

@@

Tables below shows the top 20 keywords in Indian related texts comparing to other context, respectively sorted by PMI (@tbl-kwtable_PMI) and G_signed (@tbl-kwtable_Gsigned).

```{r}
#| warning: false
#| echo: false
#| label: tbl-kwtable_PMI
#| tbl-cap-location: top
#| tbl-cap: Top 20 Keywords on Indian Related Text in Natural History (sorted by PMI)

top_scores_kw %>% 
  as_tibble() %>%
  select(type, a, PMI, G_signed) %>% 
  arrange(desc(PMI)) %>%             
  head(20) %>%                       
  kbl(col.names = c("Type", "Frequency", "PMI", r"(Signed $G^2$)")) %>% 
  kable_minimal() %>% 
  scroll_box(height = "400px")
```

@@

```{r}
#| warning: false
#| echo: false
#| label: tbl-kwtable_Gsigned
#| tbl-cap-location: top
#| tbl-cap: Top 20 Keywords on Indian Related Text in Natural History (sorted by G_Signed)

top_scores_kw %>% # also valid for top_scores_colloc
  as_tibble() %>%
  select(type, a, PMI, G_signed) %>% # select 4 columns
  arrange(desc(G_signed)) %>%             # sort by PMI (descending) 
  head(20) %>%                       # select top 30 rows
  kbl(col.names = c("Type", "Frequency", "PMI", r"(Signed $G^2$)")) %>% 
  kable_minimal() %>% 
  scroll_box(height = "400px")
```
:::

@tbl-kwPMI below shows the explanation of the top 20 keywords in Indian related text.

```{r}
#| warning: false
#| label: tbl-kwPMI

keyword_PMI_tibble <- tibble(
  Word = c("ganges", "beryls", "ichthyophagi", "megasthenes", "obsidian", "bdellium", "agates", "callaina", "condensation", "gerra", "jomanes", "nonius", "prasii", "alia", "carnelian", "cophes", "hypasis", "merchandize", "peppertree", "sacae"),
  Type = c("Proper Noun", "Noun", "Plural Noun", "Proper Noun", "Noun", "Noun", "Noun", "Noun/Proper Noun", "Noun", "Noun/Proper Noun", "Noun/Proper Noun", "Noun/Proper Noun", "Noun/Proper Noun", "Adverb", "Noun", "Noun/Proper Noun", "Noun/Proper Noun", "Noun", "Noun", "Noun/Proper Noun"),
  Exp = c("a major river in India", "a type of gemstone", "a group of people who primarily subsist on fish", "a Greek historian and diplomat", "a type of volcanic glass", "a fragrant resin obtained from certain trees", "a type of semiprecious gemstone", "pale green precious stone (lat)", "the process of vapor turning into a liquid state", "war (lat)", "a Roman nomen gentile, gens or 'family name'", "a Roman nomen gentile, gens or 'family name'", "prase, green coloured gem", "by another / different way / route", "a reddish-brown variety of chalcedony", "a river that rises in the ancient Paropamise range, eventually falling into the Indus river near its confluence with the Cophes river", "a river in north India", "goods or commodities", "a tree that produces peppercorns", "the easternmost nation of Elibe, situated to the south of Ilia and the north of Bern")
) %>% print()
